\section{Extensions}
\label{sec:cases}

\input{example2}

In this section, we discuss more challenging extensions. 
For cases with multiplication, division, or other more complicated functions at the return point, e.g., the \texttt{avg} case, we can model them as an \emph{uninterpreted $k$-ary function} and verify that all $k$ parameters of the uninterpreted functions remain the same no matter how the input is permuted, e.g., the \texttt{avg} program always produces the same \texttt{sum} and \texttt{cnt} for all permutation of the same input data word. This is a \emph{sound} but \emph{incomplete} procedure for verifying programs of this type. Nevertheless, it is not often that a  practical program for data analytics produces, e.g., $2q/2r$ from some input and $q/r$ for its permutation. Hence this procedure is often enough for proving commutativity for real world programs.

The case \texttt{MAD} (Mean Absolute Deviation) is a bit more involved. Beside the division operator $/$ that also occurs in the \texttt{avg} example, it uses a new iterator operation $\init$, which resets $\cur$ to the head of the input data word. The strategy to verify this program is to divide the task into two parts: (1) ensure that the value of \texttt{avg} is independent of the order of the input, (2) treat \texttt{avg} as a special variable whose initial value is nondeterministically assigned and then check if the 2nd half of the program (the part after $\nnext$) is commutative. The latter requires an extension to the SNT decision problems. Given a data word $w$, we say a data word $w'$ is a $k$-permutation of $w$ iff there exists $\sigma_k\in S_{|w|}$ such that $\sigma_k(i)=i$ for $i \in [k]$ and $w'=\sigma_k(w)$. In words, $\sigma_k$ is a permutation of length $|w|$ such that it does not change the first $k$ elements. 


 A \emph{$k$-commutativity problem} of a SNT $\Ss$ asks whether for each data word $w$ and its $k$-permutation $w'$, $\Ss(w)=\Ss(w')$. The problem can be reduced to SNT equivalence problems in a similar way to the standard commutativity problem.

\begin{proposition}\label{prop-snt-kcmm-to-eqv}
	The $k$-commutativity problem of SNTs is reduced to the equivalence problem of SNTs in exponential time. 
\end{proposition}



\begin{wrapfigure}{r}{0.4\textwidth}
	\vspace{-0.8cm}
	\lstset{language=C,
		basicstyle=\ttfamily\scriptsize}
	\begin{lstlisting}[mathescape=true]
int MAD2() {
  avg:= $\cur$;$\nnext$;
  loop{
    if($\cur$<avg){mad=mad+(avg-$\cur$);}
    else{mad=mad+($\cur$-avg);};$\nnext$;}
  ret mad/cnt;}
	\end{lstlisting}	
	\vspace{-0.4cm}
	\caption{The 2nd half of MAD}
	\label{fig:examples3}
	\vspace{-0.7cm}
\end{wrapfigure}
To verify the 2nd half of the MAD example, we model \texttt{avg} as a control variable and assign the first value of the data word to it (c.f., Figure~\ref{fig:examples3}). 
Then we can reduce the program commutativity problem to the $1$-commutativity problem of the corresponding SNT. We handle the division at the end of the program in the same way as we did for the \texttt{avg} program. The guarantee we obtain after the corresponding SNT is checked to be $1$-commutative is that the program outputs the same value for any value of \texttt{avg} and any permutation of the rest of the input list. 

The case \texttt{SD} (Standard Deviation) is more difficult to handle. The main difficulty comes from the use of multiplication in the middle of the program (instead of at the return point). In order to have a sound procedure to verify this kind of programs, we will need the support of uninterpreted $k$-ary functions also in the SNT transitions. However, this is not a trivial extension and we consider it as a future work.


